Paper Wireframe

### 2018-08-24 paper_wireframe_knit.Rmd
Added section with rainfall/ndvi curves and dep~waarr plots. Still has some model stuff below that will be removed/corrected

```{r, echo=F, warning=F, message=F}
library('lme4')
library('car')
library('ggplot2')
library('dplyr')
library('maps')
library('stargazer')
library('splines')
library('sf')
library('leaflet')
library('gridExtra')

@@

@@

@@

The percentage of grass cover does appear to affect departure date. * Not sure how believeable this is*

@@

The first stopover in West Africa is timed to catch rains

temp<-dat2[dat2$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)

qplot(data=temp, x=SO_startDOY, y=cumrf)+facet_wrap(~country)+theme_bw()+xlim(c(0,120))+xlab("Day of year")+ylab("Cumulative rainfall (mm)")
```

>>>>

```{r, echo=F, warning=F, message=F}
library('lme4')
library('car')
library('ggplot2')
library('dplyr')
library('maps')
library('stargazer')
library('splines')
library('sf')
library('leaflet')
library('gridExtra')
library('MuMIn')

@@

# read in model-ready dataset

dmod<-read.csv('C:/cuckoo_tracking/data/stopover_bestofday_2018_1daymin_recalc_spring_mig_summary_dead_attrib_modelready.csv', h=T)

dmod$tWA=dmod$DEPwestAF-dmod$departureGarr

dmod<-dmod[,c(1,2,6,22,27,28,29,37,39,52,57,63,74,85,87,92, 115)]

dmod<-na.omit(dmod)

@@

Taking a step back, it is important to if there is any environmental
control over how birds arrive into West Africa.

Despite rain generally arriving in West Africa at the same time, there
is variance of ~ 60 days in the times in which cuckoos arrive into
West Africa (with the bulk of birds arriving between day 55 and day 95)

There is however, little relationship between arrival date into West Africa
and departure date from West Africa

m1<-lmer(DEPwestAF~departureGarr+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF%in%85:120,])

ggplot(data=dmod, aes(x=departureGarr, y=DEPwestAF ))+xlab("Arrival into West Africa (DOY)")+ylab("Departure date West Africa (DOY)")+geom_point(shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=1.83, p=0.185 *3 outliers removed*', x = -Inf, y = Inf), hjust = 0, vjust = 1)

Note: if the above model includes the 3 outliers then a there is a significant relationship... * need to decide which is more biologically plausible*

```{r, warning=F, message=F, echo=F}
m1<-lmer(DEPwestAF~departureGarr+(1|ptt)+(1|year), data=dmod)
#Anova(m1, test.statistic = 'F')

newdat<-data.frame(DEPwestAF=1, departureGarr=25:100)
newdat$p1<-predict(m1, newdata=newdat, re.form=~0)

predmat<-model.matrix(DEPwestAF~departureGarr, data=newdat)
vcv<-vcov(m1)
semod<-sqrt(diag(predmat%*%vcv%*%t(predmat)))
newdat$lc<-newdat$p1-semod*1.96
newdat$uc<-newdat$p1+semod*1.96

qplot(data=dmod, y=DEPwestAF, x=departureGarr)+
  geom_line(data=newdat, aes(x=departureGarr, y=p1), colour='red')+
  geom_line(data=newdat, aes(x=departureGarr, y=lc), colour='red', linetype='dashed')+
  geom_line(data=newdat, aes(x=departureGarr, y=uc), colour='red', linetype='dashed')+
  ylab("Days spent in West Africa")+
  xlab("Arrival into West Africa (DOY)")+
  theme_bw()+geom_point(shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=10.47, p=0.0027', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

Because almost all birds depart West Africa between day 90 and day 110,
later arriving birds spend significantly shorter time in West Africa

```{r, warning=F, message=F, echo=F}
m1<-lmer(tWA~departureGarr+(1|year)+(1|ptt), data=dmod)

newdat<-data.frame(tWA=1, departureGarr=25:100)
newdat$p1<-predict(m1, newdata=newdat, re.form=~0)

predmat<-model.matrix(tWA~departureGarr, data=newdat)
vcv<-vcov(m1)
semod<-sqrt(diag(predmat%*%vcv%*%t(predmat)))
newdat$lc<-newdat$p1-semod*1.96
newdat$uc<-newdat$p1+semod*1.96

qplot(data=dmod, y=tWA, x=departureGarr)+
  geom_line(data=newdat, aes(x=departureGarr, y=p1), colour='red')+
  geom_line(data=newdat, aes(x=departureGarr, y=lc), colour='red', linetype='dashed')+
  geom_line(data=newdat, aes(x=departureGarr, y=uc), colour='red', linetype='dashed')+
  ylab("Days spent in West Africa")+
  xlab("Arrival into West Africa (DOY)")+
  theme_bw()+geom_point(shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=63.35, p=3.19e-09', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

When graphed with average rainfall and ndvi curved (across all WA stopovers) we see that early arriving birds into West Africa spend a long time there when greening is yet to/just starting whereas late arriving birds spend a short time at near-peak greenness. Interestingly,
the mean arrival date into west africa (day 71; black line) corresponds with earlier departures, which could be facilitated by concurrent peak greening and rainfall rates.

```{r, warning=F, message=F, echo=F}
temp<- dat2[- which(dat2$country %in% c('Angola', 'Central African Republic','Congo', 'Democratic Republic of the Congo', 'Gabon',  'Cameroon', 'Burkina Faso', 'Liberia')),]

temp3<-temp %>% group_by(variable) %>%
  summarise(Mcumrf=mean(tamcumrf), rfmin=mean(tamcumrf)-sd(tamcumrf), rfmax=mean(tamcumrf)+sd(tamcumrf),
   MaquaNDVI=mean(emodisNDVI, na.rm=T), ndmin=(mean(emodisNDVI, na.rm=T)-sd(emodisNDVI, na.rm=T)),
   ndmax=(mean(emodisNDVI, na.rm=T)+sd(emodisNDVI, na.rm=T)))

#lets make variable!

ggplot(data=dmod, aes(x=departureGarr, y=DEPwestAF))+
  geom_line(data=temp3,aes(x=variable, y=Mcumrf/2), size=1, colour='blue')+
  geom_line(data=temp3,aes(x=variable, y=rfmin/2), colour='grey', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=rfmax/2), colour='grey', linetype=2)+ 
  geom_line(data=temp3,aes(x=variable, y=MaquaNDVI*100), size=1, colour='dark green')+
  geom_line(data=temp3,aes(x=variable, y=ndmin*100), colour='orange', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=ndmax*100), colour='orange', linetype=2)+
  geom_point(aes(colour=tWA))+scale_y_continuous(breaks=seq(0, 130, 10), limits=c(0,130))+
  scale_x_continuous(breaks=seq(0, 130, 10))+
  scale_colour_gradientn(colours=topo.colors(8))+
  theme_bw()+labs(y="Departure from West Africa (DOY)", x="Arrival into West Africa (DOY)", colour='Days in WA')+geom_vline(aes(xintercept=71), colour='black')
```

Visualised another way with time coloured gradient.

```{r, warning=F, message=F, echo=F} 
df <- expand.grid(departureGarr=0:130,
                  DEPwestAF=75:130)

df$filly=df$DEPwestAF-df$departureGarr

df[df$filly<0 | df$filly>80,]$filly<-NA

ggplot() + 
  geom_tile(data=df, aes(x=departureGarr, y=DEPwestAF, fill=filly),alpha = 0.75) +      
  scale_fill_gradient(low='#ffeda0', high='#f03b20', na.value=NA)+
  geom_line(data=temp3,aes(x=variable, y=Mcumrf/2), size=1, colour='blue')+
  geom_line(data=temp3,aes(x=variable, y=rfmin/2), colour='grey', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=rfmax/2), colour='grey', linetype=2)+ 
  geom_line(data=temp3,aes(x=variable, y=MaquaNDVI*100), size=1, colour='dark green')+
  geom_line(data=temp3,aes(x=variable, y=ndmin*100), colour='orange', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=ndmax*100), colour='orange', linetype=2)+
  geom_point(data=dmod, aes(x=departureGarr, y=DEPwestAF), shape=1)+
  theme_bw()+scale_y_continuous(breaks=seq(0, 130, 10), limits=c(0,130))+
  scale_x_continuous(breaks=seq(0, 130, 10))+
  labs(y="Departure from West Africa (DOY)", x="Arrival into West Africa (DOY)", fill='Days in WA')+geom_vline(aes(xintercept=71), colour='black')

OK so are there any explanatory variables that can improve a model

# make two gradient metrics

dmod$ndvi_last_ang<-dmod$ndviTdiff_last*dmod$ndvidiff_last
dmod$rain_first_ang<-(dmod$monsTdiff_first+25)*dmod$raindiff2_first

m0<-lmer(DEPwestAF~(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m1<-lmer(DEPwestAF~dep_country+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m2<-lmer(DEPwestAF~firstCdepartureG+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m3<-lmer(DEPwestAF~departureGarr+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m4<-lmer(DEPwestAF~nSOdepartureG+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m5<-lmer(DEPwestAF~monsTdiff_first+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m6<-lmer(DEPwestAF~raindiff2_first+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m7<-lmer(DEPwestAF~emodisNDVI_last+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m8<-lmer(DEPwestAF~ndviTdiff_last+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m9<-lmer(DEPwestAF~ndvidiff_last+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m10<-lmer(DEPwestAF~emodisANOM_mean+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m11<-lmer(DEPwestAF~Trees_mean+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m12<-lmer(DEPwestAF~Croplands_mean+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])
m13<-lmer(DEPwestAF~d_river_mean+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])

m14<-lmer(DEPwestAF~rain_first_ang+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])

m15<-lmer(DEPwestAF~ndvi_last_ang+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF %in% 85:120,])

anova(m0,m1,m2,m3,m4,m5,m6,m7,m8,m9,m10, m11, m12,m13, m14, m15)

moddat<-dmod[dmod$DEPwestAF %in% 85:120,]

mfin<-lmer(DEPwestAF~departureGarr+nSOdepartureG+monsTdiff_first*raindiff2_first+emodisNDVI_last+ndviTdiff_last*ndvidiff_last+emodisANOM_mean+Trees_mean+Croplands_mean+d_river_mean+(1|ptt)+(1|year), data=moddat,  na.action = na.fail)

mdr<-dredge(mfin)

get.models(mdr, subset = delta < 4, method = "REML")

#qplot(data=dmod[dmod$DEPwestAF %in% 85:120,], y=DEPwestAF, x=nSOdepartureG)

#qplot(data=dmod[dmod$DEPwestAF %in% 85:120,], y=DEPwestAF, x=ndviTdiff_last)

#qplot(data=dmod[dmod$DEPwestAF %in% 85:120 & dmod$emodisANOM_mean>-0.05,], y=DEPwestAF, x=emodisANOM_mean)
```

@@

The percentage of grass cover does appear to affect departure date. *Not sure how believeable this is*

@@
